Brief notes on erosion

### 2019-04-22 docs/proposals/ErosionWhitePaper.tex
worked on white paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{wrapfigure}[24]{r}{0.7\textwidth}
\centering
\includegraphics[width=0.68\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} \em Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical erosion. The flow is
horizontal (left to right) and the 6 snapshots are evenly spaced in
time.  The local erosion rate on the bodies is proportional to the
magnitude of the vorticity which we compute in the fluid bulk (color
scale).  Erosion not only diminishes the size of the bodies but also
alters their shapes considerably. A few well-defined channels develop in
the space between bodies.}
\end{wrapfigure}

@@

In our recent publication~\cite{qua-moo2018}, we describe a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  incompressible Stokes equations are solved with a boundary integral
  equation (BIE).  By using a BIE, the complex geometries are naturally
  handled, high-order accurate discretizations are deployed, and fast
  summation methods result in optimal complexity.

  \item {\bf Interface Tracking}: Instead of tracking the $(x,y)$
  coordinates of the of the eroding interfaces, we track the {\thL}
  coordinates, where $\theta$ is a local tangent angle, and $L$ is the
  total interface length.  The governing equations of the {\thL}
  coordinates offers several advantages in the ability to numerically
  stabilize the evolving interface.

  \item {\bf Regularizations of the Interface}: Erosion results in
  numerically challenging corners of the bodies at stagnation points of
  the flow.  We use a regularization term and a smoothing operation to
  control the smoothness of the moving interface.

  \item {\bf Analysis of Single Eroding Bodies}: We validate our method
  by comparing numerical simulations with scaling laws.
\end{enumerate}

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, additional numerical tools are required to simulate more
realistic scenarios such as denser packings, transport, and
three-dimensional erosion.  Second, we will investigate physically
motivated questions regarding the geometry and flow inside an eroding
porous medium.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
\begin{wrapfigure}[22]{r}{0.4\textwidth}
\centering
\input{figs/vort1.tikz}
\caption{\label{fig:vort1} \em The vorticity a vertical slice cutting
through a single body.  There are solid walls at $y = \pm 1$, and an
eroding body at $y = (-0.5,0.5)$.  The trapezoid rule gives large errors
near the solid walls, while the Barycentric quadrature formula results
in a smooth vorticity field.}
\end{wrapfigure}

In its current formulation, our bodies are fixed in space.  However, in
reality, eroding bodies are also transported by the flow.  To allow the
bodies to move, there are two additional tools that we will develop.
The PI is experienced with simulating moving interfaces in Stokes flow.
For example, he has developed high-order adaptive time stepping methods
for simulating deformable bodies (vesicles) in a viscous
fluid~\cite{qua-bir2015b, qua-bir2014, qua-bir2016}. More recently, the
PI and his former Ph.D.~student developed time stepping methods for
rigid body transport in a viscous fluid~\cite{bys-sha-qua2018}.  This
work extends a contact algorithm, originally developed by Lu et
al.~\cite{lu-rah-zor2017}, to accuratly simulate the dynamics of
nearly-touching rigid bodies.  The contact algorithm resolves errors
introduced by time stepping, but quadrature error introduced by the
integral equation formulation must also be addressed.  A postdoc in the
PI's group is extending a Barycentric near-singular integration scheme
developed by Barnett et al.~\cite{bar-wu-vee2015} to develop quadrature
formula to compute the shear stress and vorticity.  In
Figure~\ref{fig:vort1}, the vorticity is computed along a vertical slice
that cuts through a single eroding body.  The trapezoid rule is accurate
far from the solid walls ($y = \pm 1$) and the boundary of the eroding
body ($y = \pm 0.5$), but introduces large errors near the walls and
body.  However, the Barycentric quadrature rule is uniformly accurate
throughout the entire fluid domain.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{wrapfigure}[22]{r}{0.6\textwidth}
\centering
\includegraphics[width=0.58\textwidth]{figs/50bod.pdf}
\caption{\label{fig:50bod} \em Simulation of 50 bodies eroding in Stokes
flow under the action of fluid-mechanical erosion. The flow is
horizontal (left to right) and the 6 snapshots are evenly spaced in
time.  The local erosion rate on the bodies is proportional to the
magnitude of the vorticity which we compute in the fluid bulk (color
scale).  Erosion both diminishes the size of the bodies and considerably
alters their shapes.}
\end{wrapfigure}

@@

In our recent publication~\cite{qua-moo2018}, we describe a suite of
numerical methods to simulate erosion in a viscous fluid
(Figure~\ref{fig:50bod}).  The highlights of our numerical methods are:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item {\bf Boundary Integral Equation Fluid Solver}: The
  fluid equations are solved with a boundary integral equation (BIE).
  The BIE formulation achives high-order accuracy in complex geometries
  with optimal computational complexity.

  \item {\bf Regularizations of the Interface}: Erosion results in
  numerically challenging corners of the bodies.  We use a
  regularization term and a smoothing operation to control the
  smoothness of the moving interface.

  \item {\bf Interface Tracking}: We describe the interface using
  {\thL} coordinates, where $\theta$ is a local tangent angle, and $L$
  is the total interface length.  In these coordinates, the
  stiff terms are linear and are be discretized implicitly.

  \item {\bf Validation}: We compare the numerical simulations with
  scaling laws.
\end{enumerate}

\section{Proposed Work}
Having developed and validated numerical methods for simulating viscous
erosion, there are two main directions that our research will follow.
First, {\bf additional numerical tools are required to simulate more
realistic scenarios such as denser packings, transport, and
three-dimensional erosion}.  Second, {\bf we will investigate physically
motivated questions regarding the geometry and flow inside an eroding
porous medium}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved Numerical Methods}
\begin{wrapfigure}[22]{r}{0.4\textwidth}
\centering
\input{figs/vort1.tikz}
\caption{\label{fig:vort1} \em The vorticity along a vertical slice
cutting through a single body.  There are solid walls at $y = \pm 1$,
and an eroding body at $y = (-0.5,0.5)$.  The trapezoid rule gives large
errors near the solid walls, while the Barycentric quadrature formula
results in a smooth vorticity field.}
\end{wrapfigure}

Our current formulation fixes all bodies space.  However, in reality,
the bodies are also transported by the flow.  The PI has simulated
moving interfaces in Stokes flow~\cite{qua-bir2015b, qua-bir2014,
qua-bir2016}. More recently, the PI and his former Ph.D.~student
developed time stepping methods for rigid body transport in Stokes
flow~\cite{bys-sha-qua2018}.  This work extends a contact algorithm,
originally developed by Lu et al.~\cite{lu-rah-zor2017}, to accuratly
simulate the dynamics of nearly-touching rigid bodies.  The contact
algorithm resolves errors introduced by time stepping, but quadrature
error introduced by the BIE formulation must also be addressed.  A
postdoc in the PI's group is extending a Barycentric near-singular
integration scheme developed by Barnett et al.~\cite{bar-wu-vee2015} to
develop quadrature formula to accurately compute the shear stress,
vorticity, and other physical quantities.  In Figure~\ref{fig:vort1},
the vorticity is computed along a vertical slice that cuts through a
single eroding body.  The trapezoid rule is accurate far from the solid
walls ($y = \pm 1$) and the boundary of the eroding body ($y = \pm
0.5$), but introduces large errors near the walls and body.  However,
our new Barycentric quadrature rule is uniformly accurate throughout the
entire fluid domain.
